### 2022-03-30 paper/localpsf.tex
done editing based on Noemi's comments
editing intro

The output of our method is a hierarchical matrix (H-matrix). H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank (see Figure \ref{fig:hmatrix_intro}). For more details about H-matrices, see Appendix \ref{app:h_matrix}. H-matrix methods permit us to perform matrix-vector products cheaply. Furthermore, one may use H-matrix methods to perform useful linear algebra operations involving the H-matrix that cannot easily be done using the original operator. These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion. These H-matrix methods are fast and scale well to large problems. The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification. Since we are building an H-matrix, it is natural to ask why we do not build the H-matrix directly. The reason is that \emph{classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products}.

Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}, which we do not use. These alternative methods have, for instance, been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}. One issue with this approach is that, while these methods are asymptotically scalable in theory, in practice the required number of matrix vector products is large\footnote{DO CALCULATION}. Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods based on the peeling process more efficient in the future.

@@

@@

	\paragraph{Feasibility} To show that $v_*$ is feasible (i.e., $v_* \in X$), we must show that $v_* \in H^2_N(\Omega)$ and $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,\convrank$. We already showed that $v_* \in H^2_N(\Omega)$ in Theorem \ref{thm:wi_no_breakdown}. Using the the definition of $v_*$ in \eqref{eq:optimal_v} and Lemma \ref{lem:S_matrix_characterizations}, we have
	\begin{equation*}
	v_*(x_i) = \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \secondgreens_j(x_i)
	= \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \pointinteractionmatrix_{ij} = \left(\pointinteractionmatrix \bm{\secondgreenscoeff} + \constcoeff \bm{1}\right)_i.
	\end{equation*}
	The conditions $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,\convrank$ may therefore be written in matrix form as
	$
	\pointinteractionmatrix \bm{\secondgreenscoeff} + \constcoeff \bm{1} = \bm{\interpolatedvalues}
	$, and this matrix equation must hold because it is the first block row of system \eqref{eq:interpolation_linear_system} which is solved to compute $\bm{\secondgreenscoeff}$ and $\constcoeff$. So $v_* \in X$.

>>>>

The output of our method is a hierarchical matrix (H-matrix). H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high-rank (see Figure \ref{fig:hmatrix_intro} and Appendix \ref{app:h_matrix}). H-matrix methods permit us to perform matrix-vector products cheaply. Furthermore, one may use H-matrix methods to perform useful linear algebra operations involving the H-matrix that cannot easily be done using the original operator. These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion. These H-matrix methods are fast and scale well to large problems. Typically, the work and memory required to perform these operations for an $N \times N$ H-matrix scales as $O(r^2 N \log(N)^a)$, where $r$ is the largest rank of the off-diagonal blocks, and $a \in \{0,1,2,3\}$. The exact cost depends on the type of H-matrix used and operation performed; for more details see \cite{HMATRIXCOST}. The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification. However, \emph{classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free}.

@@

Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}, which we do not use here. These alternative methods have, for instance, been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}. One issue with this approach is that, while these methods are asymptotically scalable in theory, in practice the required number of matrix vector products is large\footnote{DO CALCULATION}. Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods based on the peeling process more efficient in the future.

@@

	\paragraph{Feasibility} To show that $v_*$ is feasible (i.e., $v_* \in X$), we must show that $v_* \in H^2_N(\Omega)$ and $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,\convrank$. We already showed that $v_* \in H^2_N(\Omega)$ in Theorem \ref{thm:wi_no_breakdown}. Using the definition of $v_*$ in \eqref{eq:optimal_v} and Lemma \ref{lem:S_matrix_characterizations}, we have
	\begin{equation*}
	v_*(x_i) = \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \secondgreens_j(x_i)
	= \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \pointinteractionmatrix_{ij} = \left(\pointinteractionmatrix \bm{\secondgreenscoeff} + \constcoeff \bm{1}\right)_i.
	\end{equation*}
	The conditions $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,\convrank$ may therefore be written in matrix form as
	$
	\pointinteractionmatrix \bm{\secondgreenscoeff} + \constcoeff \bm{1} = \bm{\interpolatedvalues}
	$, and this matrix equation must hold because it is the first block row of system \eqref{eq:interpolation_linear_system} which is solved to compute $\bm{\secondgreenscoeff}$ and $\constcoeff$. So $v_* \in X$.
